Forecasting biodiversity in breeding birds using best practices

### 2017-12-08 manuscript/draft.Rmd
is this unfair to SDMs?
in response to reviewer 2

Forecasting is an emerging imperative in ecology; as such, the field needs to
develop and follow best practices for conducting and evaluating ecological
forecasts [@clark2001]. We have used a number of these practices (Box 1) in a
single study that builds and evaluates forecasts of biodiversity in the form of
species richness. The results of this effort are both promising and humbling.
When comparing forecasts across sites, many different approaches to forecasting
produce reasonable forecasts (Figure 3). If a site is predicted to have a high
number of species in the future, relative to other sites, it generally does.
However, none of the methods evaluated reliably determined how site-level
richness changes over time (Figure 6), which is generally the stated purpose of
these forecasts. As a result, baseline models, which did not attempt to
anticipate changes in richness over time, generally provided the best forecasts
for future biodiversity. While this study is restricted to breeding birds in
North America, its results are consistent with a growing literature on the
limits of ecological forecasting, as discussed below.

The most commonly used methods for forecasting future biodiversity, SDMs and
macroecological models, both produced worse forecasts than time-series models
and simple baselines. This weakness suggests that predictions about future
biodiversity change should be viewed with skepticism unless the underlying
models have been validated temporally, via hindcasting and comparison with
simple baselines. Since site-level richness is relatively stable, spatial
validation is not enough: a model can have high accuracy across spatial
gradients without being able to predict changes over time. This gap between
spatial and temporal accuracy is known to be important for species-level
predictions [@rapacciuolo2012; @oedekoven2017]; our results indicate that it is
substantial for higher-level patterns like richness as well. SDMs' poor temporal
predictions are particularly sobering, as these models have been one of the main
foundations for estimates of the predicted loss of biodiversity to climate
change over the past decade or so [@thomas2004; @thuiller2011; @urban2015]. Our
results also highlight the importance of comparing multiple modeling approaches
when conducting ecological forecasts, and in particular, the value of comparing
results to simple baselines to avoid over-interpreting the information present
in these forecasts [Box 1]. Disciplines that have more mature forecasting
cultures often do this by reporting "forecast skill", i.e., the improvement in
the forecast relative to a simple baseline [@jolliffe2003]. We recommend
following the example of @ye2015 and adopting this approach in future ecological
forecasting research.

>>>>

Forecasting is an emerging imperative in ecology; as such, the field needs to
develop and follow best practices for conducting and evaluating ecological
forecasts [@clark2001]. We have used a number of these practices (Box 1) in a
single study that builds and evaluates forecasts of biodiversity in the form of
species richness. The results of this effort are both promising and humbling.
When comparing forecasts across sites, many different approaches to forecasting
produce reasonable forecasts (Figure 3). If a site is predicted to have a high
number of species in the future, relative to other sites, it generally does.
However, none of the methods evaluated reliably determined how site-level
richness changes over time (Figure 6), which is generally the stated purpose of
these forecasts. As a result, baseline models, which did not attempt to
anticipate changes in richness over time, generally provided the best forecasts
for future biodiversity. While this study is restricted to breeding birds in
simple baselines. Since site-level richness is relatively stable in so many
systems [@dornelas2014], spatial validation will not be enough: a model can have
high accuracy across spatial gradients without being able to predict changes
over time. This gap between spatial and temporal accuracy is known to be
important for species-level predictions [@rapacciuolo2012; @oedekoven2017]; our
results indicate that it is substantial for higher-level patterns like richness
as well. SDMs' poor temporal predictions are particularly sobering, as these
models have been one of the main foundations for estimates of the predicted loss
of biodiversity to climate change over the past decade or so [@thomas2004;
@thuiller2011; @urban2015]. One could argue that these conclusions are unfair to
SDMs and related methods, which are generally used to predict large directional
changes over multi-decade timescales rather than fluctuations within a single
decade. It is certainly possible that these models' relative performance would
improve over time or with larger shifts in richness, and future research should
explore this possibility.

Our results also highlight the importance of comparing multiple modeling
approaches when conducting ecological forecasts, and in particular, the value of
comparing results to simple baselines to avoid over-interpreting the information
present in these forecasts [Box 1]. Disciplines that have more mature
forecasting cultures often do this by reporting "forecast skill", i.e., the
improvement in the forecast relative to a simple baseline [@jolliffe2003]. We
recommend following the example of @ye2015 and adopting this approach in future
ecological forecasting research.
